# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
The data file (activity.zip) was included in the forked repo.  Since it doesn't have to be downloaded it's simple to unzip it and read it into the variable `act`:

```{r initialization, results='hide', message=FALSE}
###### Initialization
## Clear the workspace
rm(list = ls(all = TRUE))

## Load required libraries
stopifnot(require(dplyr))
stopifnot(require(ggplot2))

## Read in the data and massage it to the desired formats/factors
zipName <- "./activity.zip"
unzip(zipName)
fileName <- "./activity.csv"
act <- read.csv(fileName, 
                header = TRUE, 
                na.strings = "NA", 
                colClasses = c("integer", "character", "integer"))
act$date <- as.Date(act$date, "%Y-%m-%d")
act$date <- as.factor(act$date)
```

## What is mean total number of steps taken per day?
I find that the _dplyr_ package makes it very simply to calculate aggregates across subsets.  In this case we first use the `summarise()` function to calculate the total number of steps taken per day, then take the mean and median of those steps:
```{r totalSteps, message=FALSE}
###### Total number of steps per day
## get the total per day and plot it as a histogram
act.no_na <- na.omit(act)
steps.total <- summarise(group_by(act.no_na,date), sum(steps))
names(steps.total)[2] <- "total"
qplot(steps.total[, 2], 
      data = steps.total, 
      geom="histogram", 
      xlab="Steps", 
      ylab="Count",
      main="Total Steps per Day")
```

The mean number of steps per day (ignoring NAs) is:
```{r}
mean(steps.total[, 2])
```

and the median number of steps per day (ignoring NAs) is:
```{r}
median(steps.total[, 2])
```

## What is the average daily activity pattern?
```{r}
steps.interval <- summarise(group_by(act.no_na, interval), mean(steps))
names(steps.interval)[2] <- "total"
ggplot(steps.interval, aes(interval, total)) +
  geom_line() +
  ylab("total number of steps") +
  ggtitle("Steps per 5 Minute Interval")
```

The interval with the most steps is:
```{r}
max(steps.interval[, 2])  # max number of steps in an interval
#si.max  # print it out
#si.interval <- steps.interval[steps.interval$total == si.max] # interval with that max
```

## Inputing missing values



## Are there differences in activity patterns between weekdays and weekends?
